Low Complexity Adaptive Beamformer\\ for Active Sonar Imaging

### 2015-10-04 LCAforActiveSonarImaging/lca.tex
la2pdf commit - 2015:10:04	22:10:59	(Sun 4.Oct)

% 
% \begin{itemize}
% \item Dynamic weights $\Rightarrow$ greater contrast/resolution potential
% \item But:
% \begin{itemize}
% \item Active system $\Rightarrow$ noise and signal correlated $\Rightarrow$ adaptive beamformers no longer robust.
% \item Subarray averaging decorrelate the noise, but in the process we sacrifice resolution.
% \item Computationally complex O($M^3$) due to inversion of a spatial covariance matrix. LCA is only O($MW$). Even if Capon subspace/beamspace versions exist, they remain more complex than the LCA, and is not as ideally suited for implementation of SIMD hardware (GPUs, DSPs, ...).
% \end{itemize}
% \item LCA is inherently robust, is ``parameter free'' once setup with a well designed window set, and can often be implemented with minor modifications to already existing DAS beamformers.
% 
% The angular resolution and contrast in active sonar images depend on the beamformer's ability to receive signals from directions of interest, while suppressing noise and interference emanating from other directions. For sonar arrays, this is achieved by applying weights to the array channels.
% 
% While classical beamformers use predefined windows, adaptive beamformers estimate the optimal window by analytical evaluation of the data. The minimum variance (MV) beamformer, for instance, calculates the set of weights that minimises the variance of the beamformer's output. 
% 
We have implemented a Low Complexity Adaptive (LCA) beamformer, which adaptively selects a window from a predefined set. The set is comprised of windows that are typical solutions found by the Minimum Variance method. The LCA beamformer was tested using simulated and experimental data from the Kongsberg Maritime HISAS 1030 sonar. On a simulated scene with speckle, highlight and shadow, the beamformer offered better lateral edge definition compared to the MV beamformer, and speckle intensity and shape comparable to DAS and MV beamformers. These results were verified by the experimental data.
% 
% An attactive trait of the LCA is its low computational complexity; while the MV beamformer is of O($M^3$), the LCA method is of O($MW$), with $M$ being the number of channels and $W$ the number of windows. We made the LCA perform like the MV method using a well designed set of 30 windows. Hence, unless the array is very small, the proposed method will perform like the MV beamformer or better, and at a fraction of the computational cost.

@@

This method is called the Low Complexity Adaptive (LCA) beamformer. It was first proposed by Synnev\aa{}g et. al. in clinical medical imaging~\cite{Synnevag2008}, who demonstrated its ability to obtain very similar results to MVDR. It does this by applying a set of predefined windows and selects the one that is best able to suppress noise. Since it uses the same optimization criterion as MVDR, it may be considered a version of MVDR with a reduced window solution space. As we will demonstrate, this works well because a robust MVDR design tends to create windows that resembles slightly varied versions of conventional ones. But which windows should LCA choose from?

Synnev\aa{}g et. al. used a window set comprised of rectangular, Kaiser and inverted Kaiser functions\footnote{May also be called Kaiser-Harris windows.}. Of these a specific Kaiser window were micro-steered. His database had 12 windows in it, which proved to work well. But just how well did this work compared to MVDR? Can LCA perform even better with more windows, and could it work just as well with even fewer?

@@

Our findings show that when applied to experimental data, the weights chosen by a robustified MVDR yield responses that tend to be symmetric and steered within a fraction of the -3dB resolution of the respective window. When letting LCA choose from Kaiser windows with similar responses, we obtain images with both improved noise suppression and resolution. The resolution gain can be made predictable and depending on the amount of steering we apply to the windows. Overall, the image quality produced by LCA is comparable to MVDR, and sometimes even better. LCA is also inherently robust, easy to implement and fairly easy to understand.

>>>>

% 
% \begin{itemize}
% \item Dynamic weights $\Rightarrow$ greater contrast/resolution potential
% \item But:
% \begin{itemize}
% \item Active system $\Rightarrow$ noise and signal correlated $\Rightarrow$ adaptive beamformers no longer robust.
% \item Subarray averaging decorrelate the noise, but in the process we sacrifice resolution.
% \item Computationally complex O($M^3$) due to inversion of a spatial covariance matrix. LCA is only O($MW$). Even if Capon subspace/beamspace versions exist, they remain more complex than the LCA, and is not as ideally suited for implementation of SIMD hardware (GPUs, DSPs, ...).
% \end{itemize}
% \item LCA is inherently robust, is ``parameter free'' once setup with a well designed window set, and can often be implemented with minor modifications to already existing DAS beamformers.
% 
% The angular resolution and contrast in active sonar images depend on the beamformer's ability to receive signals from directions of interest, while suppressing noise and interference emanating from other directions. For sonar arrays, this is achieved by applying weights to the array channels.
% 
% While classical beamformers use predefined windows, adaptive beamformers estimate the optimal window by analytical evaluation of the data. The minimum variance (MV) beamformer, for instance, calculates the set of weights that minimises the variance of the beamformer's output. 
% 
We have implemented and studied the low complexity adaptive (LCA) beamformer in active sonar imaging. LCA is an attractive alternative to the minimum variance distortionless response (MVDR) method for minimizing noise and interference in sonar images. While MVDR estimates and inverts a spatial covariance matrix to compute an optimal window for the sonar array, LCA selects the optimal window from a predefined set using the same criterion as MVDR.

We show that a robust MVDR implementation typically chooses approximately symmetric windows with maximum sidelobe levels between that of the rectangular and Hamming window. We then set up LCA to use Kaiser windows with similar responses to those of MVDR, and apply various amounts of steering. We find that when Steering to -3\,dB width of the window. This lead to a predictable resolution gain of . design LCA's window database with windows that resembles  constrained to work in active sonar imaging the typical window responses LCA is allowed to choose from windows that are typical solutions found by the minimum variance distortionless response (MVDR) method.

The LCA beamformer was tested using simulated and experimental data from the Kongsberg Maritime HISAS 1030 sonar. On a simulated scene with speckle, highlight and shadow, the beamformer offered better lateral edge definition compared to the MV beamformer, and speckle intensity and shape comparable to DAS and MV beamformers. These results were verified by the experimental data.
% 
% An attactive trait of the LCA is its low computational complexity; while the MV beamformer is of O($M^3$), the LCA method is of O($MW$), with $M$ being the number of channels and $W$ the number of windows. We made the LCA perform like the MV method using a well designed set of 30 windows. Hence, unless the array is very small, the proposed method will perform like the MV beamformer or better, and at a fraction of the computational cost.

@@

This method is called the Low Complexity Adaptive (LCA) beamformer. It was first proposed by Synnev\aa{}g in clinical medical imaging~\cite{Synnevag2008}, who demonstrated its ability to obtain very similar results to MVDR. It does this by applying a set of predefined windows and selects the one provide the best noise suppression. It uses the same optimization criterion as MVDR, so it may be considered a version of MVDR with a reduced window solution space. As we will demonstrate, this works well because a robust MVDR design tends to create windows that resembles slightly varied versions of conventional ones. But which windows should LCA choose from?

Synnev\aa{}g used a window set comprised of rectangular, Kaiser and inverted Kaiser functions\footnote{May also be called Kaiser-Harris windows.}. Of these a specific Kaiser window were micro-steered. His database had 12 windows in it, which proved to work well. But just how well did this work compared to MVDR? Can LCA perform even better with more windows, and could it work just as well with even fewer?

@@

Our findings show that when applied to experimental data, the weights chosen by a robustified MVDR yield responses that tend to be symmetric and steered within a fraction of the -3dB resolution of the respective window. When letting LCA choose from Kaiser windows with similar responses, we obtain images with improved noise suppression and resolution. The resolution gain is made predictable by limiting the maximum steering to a fraction of the window's -3\,dB width. Overall, LCA produces images that are similar to MVDR, but with a resolution in between that of DAS and MVDR. LCA is also inherently robust, easy to implement and fairly easy to understand.
